---
title: "**San Juan in-season update**"
date: 'Last update: `r Sys.Date()`'
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message=F)


# Load packages ---------------------------------
library(tidyverse)


# Helpers --------------------------------- <<<UPDATE>>>
escapement_year <- 2024


# ================== COUNT DATA ================== 
# SIL data ---------------------------------
SJ.SIL <- readxl::read_excel(path=paste0("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/", 
                                               escapement_year, 
                                               "/SILs/Area 20/San Juan River/A20_SanJuan_River_", 
                                               escapement_year, 
                                               ".xlsx"),
                                   sheet="San Juan raw SIL entry", guess_max=1000) %>% 
   mutate_if(is.numeric, ~replace_na(., 0)) %>% 
   mutate_at(c("adults # holding", "adults live / habitat seen (# Est Live)"), ~round(.,0)) %>% 
   mutate(date = lubridate::ymd(date)) %>%
   print()


# ================== ENVIRO DATA ================== 
# Water data ---------------------------------
SJ.water <- full_join(
  read.csv(file=list.files(path = here::here("data", "enviro"),
                           pattern = "SANJUAN_historical*",
                           full.names = TRUE),
           skip=1) %>%
    mutate(PARAM = case_when(PARAM==1 ~ "Discharge (cms)",
                             PARAM==2 ~ "Level (m)"),
           Date = lubridate::ymd(Date),
           year = lubridate::year(Date),
           month = lubridate::month(Date, label=T, abbr=T),
           DOY = lubridate::yday(Date)),
  
  rbind(read.csv(file=list.files(path = here::here("data", "enviro"),
                                 pattern = "SANJUAN_realtime_08HA010_QR*",     # discharge realtime
                                 full.names = TRUE),
                 skip=9) %>%
          rename(Value = Value..m..s.,
                 Date = Date..PST.,
                 PARAM = Parameter),
        
        read.csv(file=list.files(path = here::here("data", "enviro"),
                                 pattern = "SANJUAN_realtime_08HA010_HG*",     # water level realtime
                                 full.names = TRUE),
                 skip=9) %>%
          rename(Value = Value..m.,
                 Date = Date..PST.,
                 PARAM = Parameter)
  )
  %>%
    mutate(PARAM = case_when(PARAM==46 ~ "Level (m)",
                                 PARAM=="47" ~ "Discharge (cms)"),
           Date = lubridate::ymd(stringr::str_sub(string=Date, start=1, end=10)),
           time = stringr::str_sub(string=Date, start=12, end=19),
           year = lubridate::year(Date),
           month = lubridate::month(Date, label=T, abbr=T),
           DOY = lubridate::yday(Date)) 
)



# Nitinat water data for comparison ---------------------------------
nitinat.water <- 
  # lapply(list.files(path = here::here("data", "enviro"),
  #                   pattern = "^NitinatWaterLevelHydrometData.*\\.xlsx$",
  #                   full.names = TRUE), 
  #        readxl::read_excel, sheet="Exported Data") %>% 
  #do.call("cbind",.) %>%
  lapply(list.files(path = here::here("data", "enviro"),
                    pattern = "NitinatWaterLevelHydrometData_20*",
                    full.names = TRUE),
         read.csv, skip=1) %>%
  do.call("rbind",.) %>%
  separate(Received, sep=" ", c("date", "time", "stamp")) %>%
  mutate(date2 = case_when(is.na(TL.Battery.V) ~ lubridate::mdy(date),
                          !is.na(TL.Battery.V) ~ lubridate::dmy(date)),
        year = lubridate::year(date2),
        month = lubridate::month(date2, label=T, abbr=T),
        DOY = lubridate::yday(date2),
        Water.Level.m = case_when(grepl("\\*", Water.Level.m) ~ NA,
                                  TRUE ~ as.numeric(Water.Level.m)),
        label_group = case_when(year<escapement_year ~ "Historical",
                                        year==escapement_year ~ as.character(year)),
        hour = as.numeric(stringr::str_sub(stringr::str_pad(time, width=5, pad = "0"), 1,2)),
        hour = case_when(hour==12 & stamp=="AM" ~ 0,
                         TRUE ~ hour)) %>%
  print()


# ================== HISTORICAL DATA ================== 
# Run timing ---------------------------------
sj.rtStWk <- readxl::read_excel(path=#paste0(here::here("data", "escapement"), sep="/", 
                                      #       list.files(path=here::here("data", "escapement"),
                                      #                  pattern="run timing tables (updated)*")), 
                                   "//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/run timing tables (updated).xlsx",
                                 sheet="San Juan", guess_max=10000, trim_ws=T, 
                                 range=cellranger::cell_limits(c(4, 1), c(NA, NA))) %>% 
  select(-c(`...25`:NOTES, `...34`:`...43`)) %>%
  rename(year=`...1`) %>% 
  filter(!is.na(year), !grepl("Observed", year), year!="Year") %>% 
  mutate(across(c(`81`:`124`), ~as.numeric(.)),
         plot_group = case_when(year==escapement_year ~ year,
                                year!=escapement_year ~ paste0(min(year), 
                                                               sep="-", (escapement_year-1)))) %>%
  pivot_longer(cols = c(`81`:`124`), names_to = "stat_week", values_to = "count") %>% 
  print()

sj.rtStWk$stat_week <- factor(sj.rtStWk$stat_week,
                               levels=c("81","82","83","84",
                                        "91","92","93","94",
                                        "101","102","103","104","105",
                                        "111","112","113","114","115",
                                        "121","122","123","124", ordered=T)) 
```

<br>


<br>

# **Water level and temperature**

The following water data are compiled from the San Juan Hydromet operated by Environment Canada: 

```{r fig.cap='San Juan hydromet water level (m). Historical shaded area represents daily maximum and minimum recorded values over the time series, while the line represents the mean water level. Current year line represents daily mean water level (blue). Water level is recorded every minute at the Environment Canada hydromet in the lower San Juan (above Fairy Lake)', fig.width=8, fig.height=6}

pdf(file = here::here("outputs", "figures", "San Juan water level 2024.pdf"),   
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

sjlevel <- ggplot()+
  geom_ribbon(data=SJ.water %>% 
                filter(month %in% c("Aug","Sep","Oct", "Nov") & year!=escapement_year & PARAM=="Level (m)") %>%
                group_by(DOY) %>% 
                summarize(max=max(Value,na.rm=T), min=min(Value,na.rm=T)),
              aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min, ymax=max), fill="gray80") +
  geomtextpath:: geom_textline(data=SJ.water %>% 
                                 filter(month %in% c("Aug","Sep","Oct", "Nov") & year!=escapement_year & PARAM=="Level (m)") %>%
                                 group_by(DOY) %>% 
                                 summarize(mean_hist=mean(Value,na.rm=T)) %>%
                                 mutate(label = paste0(min(SJ.water$year), sep="-", escapement_year-1)),
                              aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_hist, label=as.factor(label)), 
                              colour="gray40", linewidth=1, show.legend=F, hjust=0.92, text_smoothing=30) +
  geomtextpath:: geom_labelline(data=SJ.water %>% 
                                filter(month %in% c("Aug","Sep","Oct", "Nov") & year==escapement_year & PARAM=="Level (m)") %>%
                                group_by(DOY) %>% 
                                summarize(mean=mean(Value,na.rm=T)) %>% 
                                mutate(label=escapement_year),
                              aes(x=as.Date(DOY, origin="2021-12-31"), y=mean, label=as.factor(label)), 
                              colour="dodger blue", linewidth=1, hjust=0.05, show.legend=F) +
  labs(y="Water level (m)") +
  scale_colour_manual(values=c("dodger blue", "gray70")) +
  scale_x_date(date_labels = "%b %d") +
  theme_bw() +
  theme(axis.text = element_text(colour="black"),
        axis.title = element_text(face="bold"),
        axis.title.x = element_blank(),
        legend.title = element_blank())

dev.off()

sjlevel
```

```{r fig.cap='San Juan hydromet discharge (cms). Historical shaded area represents daily maximum and minimum recorded values over the time series, while the line represents the mean water level. Current year line represents daily mean water level (blue). Disharge is recorded every minute at the Environment Canada hydromet in the lower San Juan (above Fairy Lake).', fig.width=8, fig.height=6}

pdf(file = here::here("outputs", "figures", "San Juan water discharge 2024.pdf"),   
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

sjflow <- ggplot()+
  geom_ribbon(data=SJ.water %>% 
                filter(month %in% c("Aug","Sep","Oct", "Nov") & year!=escapement_year & PARAM=="Discharge (cms)") %>%
                group_by(DOY) %>% 
                summarize(max=max(Value,na.rm=T), min=min(Value,na.rm=T)),
              aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min, ymax=max), fill="gray80") +
  geomtextpath:: geom_textline(data=SJ.water %>% 
                                 filter(month %in% c("Aug","Sep","Oct", "Nov") & year!=escapement_year & PARAM=="Discharge (cms)") %>%
                                 group_by(DOY) %>% 
                                 summarize(mean_hist=mean(Value, na.rm=T)) %>%
                                 mutate(label = paste0(min(SJ.water$year), sep="-", escapement_year-1)),
                              aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_hist, label=as.factor(label)), 
                              colour="gray40", linewidth=1, show.legend=F, hjust=0.55, text_smoothing=30) +
  geomtextpath:: geom_labelline(data=SJ.water %>% 
                                filter(month %in% c("Aug","Sep","Oct", "Nov") & year==escapement_year & PARAM=="Discharge (cms)") %>%
                                group_by(DOY) %>% 
                                summarize(mean=mean(Value,na.rm=T)) %>% 
                                mutate(label=escapement_year),
                              aes(x=as.Date(DOY, origin="2021-12-31"), y=mean, label=as.factor(label)), 
                              colour="dodger blue", linewidth=1, hjust=0.05, show.legend=F) +
  labs(y="Discharge (cms)") +
  scale_colour_manual(values=c("dodger blue", "gray70")) +
  scale_x_date(date_labels = "%b %d") +
  theme_bw() +
  theme(axis.text = element_text(colour="black"),
        axis.title = element_text(face="bold"),
        axis.title.x = element_blank(),
        legend.title = element_blank())

dev.off()

sjflow
```

<br>

<br>

<br>

## Environment Canada river forecast 

San Juan River Environment Canada COFFEE forecast is produced daily by the River Forecast Center. In a pinch, Nitinat relative levels can also be viewed in the event the San Juan gauge goes down. 

```{r}
knitr::include_url("http://bcrfc.env.gov.bc.ca/fallfloods/coffee/08HA010.PDF")
```

<br>

```{r fig.cap='Summer/fall water level at Nitinat (blue) and San Juan (gray) from 2021-present. Shaded region represents daily maximum and minimum levels while solid line indicates daily average among all years. San Juan water level from Environment Canada, Nitinat from South Coast Stock assessment hydromet.'}

ggplot()+
  geom_ribbon(data=nitinat.water %>% 
                filter(month %in% c("Aug","Sep","Oct", "Nov") & Water.Level.m>0.1) %>%
                group_by(DOY) %>% 
                summarize(max_lvl=max(Water.Level.m,na.rm=T), min_lvl=min(Water.Level.m,na.rm=T)),
              aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min_lvl, ymax=max_lvl), fill="gray70", alpha=0.5) +
  geomtextpath:: geom_textline(data=nitinat.water %>% 
                                 filter(month %in% c("Aug","Sep","Oct", "Nov") & Water.Level.m>0.1) %>%
                                 group_by(DOY) %>% 
                                 summarize(mean_lvl=mean(Water.Level.m,na.rm=T)) %>% 
                                 mutate(label = paste0("Nitinat ", min(nitinat.water$year), sep="-", escapement_year)),
                               aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_lvl, label=as.factor(label)), 
                               colour="gray70", linewidth=1, show.legend=F, hjust=0.01, alpha=0.5) +
  geom_ribbon(data=SJ.water %>% 
                filter(month %in% c("Aug","Sep","Oct", "Nov") & year%in%c(2021:escapement_year) & PARAM=="Level (m)")  %>%
                group_by(DOY) %>% 
                summarize(max_lvl=max(Value,na.rm=T), min_lvl=min(Value,na.rm=T)),
              aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min_lvl, ymax=max_lvl), fill="dodger blue", alpha=0.6) +
  geomtextpath::geom_textline(data=SJ.water %>% 
                                filter(month %in% c("Aug","Sep","Oct", "Nov") & year%in%c(2021:escapement_year) & 
                                         PARAM=="Level (m)")  %>%
                                group_by(DOY) %>% 
                                summarize(mean_lvl_hist=mean(Value,na.rm=T)) %>%
                                mutate(label = paste0("San Juan ", min(nitinat.water$year), sep="-", escapement_year)),
                              aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_lvl_hist, label=as.factor(label)), 
                              colour="dodger blue", linewidth=1, show.legend=F, hjust=0, text_smoothing=30, alpha=0.4) +
  labs(y="Water level (m)") +
  scale_colour_manual(values=c("dodger blue", "gray70")) +
  scale_x_date(date_labels = "%b %d") +
  coord_cartesian(ylim=c(0.75, 8)) +
  theme_bw() +
  theme(axis.text = element_text(colour="black"),
        axis.title = element_text(face="bold"),
        axis.title.x = element_blank(),
        legend.title = element_blank())
```

<br>

<br>

# **Live and Dead Totals** 

<br> 

The `r escapement_year` San Juan River and tribs surveys will begin on September 4. A preliminary swim was conducted prior to this to confirm Sockeye presence. See \@ref(tab:survey-table) for segments surveyed each date. 

<br>

```{r survey-table, tab.cap='Segments surveyed by date.'}
SJ.SIL %>% 
  filter(!is.na(date)) %>% 
  group_by(date, `upper/lower`) %>% 
  summarize() %>% 
  rename(`Segment(s) surveyed`=`upper/lower`,
         Date=date) %>% 
  kableExtra::kbl(align="c") %>%
  kableExtra::kable_paper("hover", full_width=T, position="center") 
```

<br>

<br>

```{r fig.cap='Raw live (bar) and carcass (line) counts for Nitinat River. Not all surveys are full system: early surveys are lower river only. Dead counts will populate once die-off begins.Note differing y-axis for each species.'}
ggplot() +
  # ---- DEAD -----
  geom_point(data=SJ.SIL %>%
               filter(!grepl("No fish", species, ignore.case=T) & !is.na(species)) %>%
               group_by(date, species) %>%
               summarize(total_adults_ded=sum(`adults dead`)) ,
             aes(x=date, y=total_adults_ded, group=species, colour=species)) +
  geom_line(data=SJ.SIL %>%
              filter(!grepl("No fish", species, ignore.case=T) & !is.na(species)) %>%
              group_by(date, species) %>%
              summarize(total_adults_ded=sum(`adults dead`)),
            aes(x=date, y=total_adults_ded, group=species, colour=species)) +
  # ---- LIVE ----
  geom_bar(data=SJ.SIL %>% 
             filter(!grepl("No fish", species, ignore.case=T) & !is.na(species)) %>% 
             group_by(date, species) %>% 
             summarize(total_adults=sum(`adults live`)), 
           aes(x=date, y=total_adults, group=species, fill=species), stat="identity", alpha=0.8, width=2) +
  facet_wrap(~species, scales="free_y") +
  #scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
  #scale_y_continuous(sec.axis = sec_axis(~ . /100, name = "Total dead adults (line)")) +
  labs(x="", y="Total live adults (bar)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position = "none", 
        axis.title = element_text(face="bold"))
```

<br>

<br>

Total counts:

```{r}
SJ.SIL %>% 
  filter(!is.na(date), !grepl("no fish", species, ignore.case=T)) %>% 
  group_by(date, species) %>% 
  summarize(adults_live = sum(`adults live`)) %>% 
  kableExtra::kbl(align="c") %>%
  kableExtra::kable_paper("hover", full_width=T, position="center")
```

Peak counts:

```{r}
SJ.SIL %>% 
  filter(!is.na(date), !grepl("no fish", species, ignore.case=T)) %>% 
  group_by(date, species) %>% 
  summarize(adults_live = sum(`adults live`)) %>% 
  group_by(species) %>% 
  filter(adults_live==max(adults_live)) %>%
  kableExtra::kbl(align="c") %>%
  kableExtra::kable_paper("hover", full_width=T, position="center")
```

<br>

<br>

```{r fig.cap='Live counts (bars) and water level (blue line) in the Nitinat River.'}
# Live ~ water 
# Plot counts with hydromet data to show pulses of fish into the system relative to rain events. 

ggplot() +
  # ---- WATER DATA ----
  geom_ribbon(data=SJ.water %>% 
                filter(year==escapement_year, PARAM=="Level (m)", Date>=("2024-09-15")) %>%
                group_by(Date) %>% 
                summarize(mean_water = mean(Value), min=min(Value), max=max(Value)),
              aes(x=Date, ymin=min*100, ymax=max*100), fill="dodger blue", alpha=0.5) +
  geom_line(data=SJ.water %>% 
              filter(year==escapement_year, PARAM=="Level (m)", Date>=("2024-09-15")) %>%
              group_by(Date) %>% 
              summarize(mean_water = mean(Value), min=min(Value), max=max(Value)),
            aes(x=Date, y=mean_water*100), colour="dodger blue", size=1) +
  # ---- COUNTS ----
  geom_bar(data=SJ.SIL %>% 
             filter(!grepl("No fish", species, ignore.case=T), !is.na(species)) %>% 
             group_by(date, species) %>% 
             summarize(total_adults=sum(`adults live / habitat seen (# Est Live)`)), 
           aes(x=date, y=total_adults, group=species, fill=species, colour=species), 
           stat="identity", alpha=0.8, width=2, size=1) +
  scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
  scale_y_continuous(sec.axis = sec_axis(~./100, name = "water level (m)")) +
  labs(x="", y="Live count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position = "none",
        axis.title = element_text(face="bold")) +
  facet_wrap(~species, ncol=2, scales="free_y")
```

<br>

<br>

# **Historical run timing** 

```{r fig.cap='Historical run timing for focal species in Nitinat River (gray). Stat weeks along x axis denote the month and week of historical counts (e.g., stat week 81=First week of August, statweek 93=Third week of Sept, etc). Current year counts will show as blue points when populated.', fig.width=8, fig.height=6}

pdf(file = here::here("outputs", "figures", "San Juan run timing.pdf"),   
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

rtall <- ggplot() +
  # annotate("text", x=Inf, y=-Inf, label="NOT FINAL ESTIMATES. DO NOT SHARE.",
  #          hjust=1.1, vjust=-1.1, col="gray70", size=6, fontface="bold", alpha=0.6) +
  geom_point(data=sj.rtStWk %>%
               group_by(species_R, year, stat_week) %>%
               summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                         is.na(count) ~ NA), plot_group=plot_group) %>%
               filter(!is.na(count)),
             aes(x=stat_week, y=count, colour=plot_group, fill=plot_group, size=plot_group),
             shape=21, alpha=0.6, stroke=1.5) +
  scale_fill_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_colour_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_size_manual(breaks=waiver(), values=c(2, 4)) +
  scale_y_continuous(labels=scales::comma) +
  labs(x="Stat week", y="Live adults (raw)", fill="Year:", colour="Year:", size="Year:", 
       #caption="Historical run timing for focal species in San Juan River (gray). Stat weeks along x axis denote the month and week of historical counts (e.g., stat week 81=First week of August, \nstatweek 93=Third week of Sept, etc). Current year counts will show as blue points when populated. DFO South Coast Salmon Stock Assessment data."
       ) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "horizontal",
        axis.title = element_text(face="bold"),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)) +
  facet_wrap(~stringr::str_to_title(species_R), nrow=3, scales="free_y") 

dev.off()

rtall
```

<br>

<br>

```{r fig.cap='Historical run timing for focal species in Nitinat River (gray). Stat weeks along x axis denote the month and week of historical counts (e.g., stat week 81=First week of August, statweek 93=Third week of Sept, etc). Current year counts will show as blue points when populated.', fig.width=8, fig.height=6}

pdf(file = here::here("outputs", "figures", "San Juan run timing (heatmap).pdf"),   
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

rtheat <- cowplot::plot_grid(

ggplot() +
  geom_tile(data=sj.rtStWk %>%
              filter(species_R=="Chinook") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C") +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Chinook\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.22),
        legend.text = element_text(size=6),
        legend.title = element_text(size=7),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold"),
        axis.text = element_text(colour="black"),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)),

ggplot() +
  geom_tile(data=sj.rtStWk %>%
              filter(species_R=="Coho") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C") +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Coho\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.22),
        legend.text = element_text(size=6),
        legend.title = element_text(size=7),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold"),
        axis.text = element_text(colour="black"),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)),

ggplot() +
  geom_tile(data=sj.rtStWk %>%
              filter(species_R=="Chum") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C", guide = guide_colorbar(barwidth = 0.5, barheight = 5)) +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Chum\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.22),
        legend.text = element_text(size=6),
        legend.title = element_text(size=7),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold"),
        axis.text = element_text(colour="black"),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)),

nrow=2)

dev.off()

rtheat
```












