---
title: "San Juan TECH REPORT DRAFT"
output: pdf_document
date: "2024-08-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)

source(here::here("reports", "source-purse_seine.R"))
```



# Acknowledgements

PFN, StA, 4 Mile, SEP (HW, MT, Melissa), DBurt

# Intro

# Methods 

## Study area and field survey methods

The San Juan River is the major salmon producing system that empties into Port San Juan (Port Renfrew, BC) on the southwest coast of Vancouver Island (MAP) 
It supports genetically distinct populations of Chinook and Coho (CITE), as well as a small abundance of SWVI Chum. The tributaries of the San Juan are particularly productive for Coho and Steelhead. 
This area falls within the traditional territory of PFN, who are committing tremendous effort towards re-establishing the natural Chinook run through restoration and monitoring projects as well as co-management of the 4 Mile Hatchery. The hatchery has enhanced chinook since x, coho since x, … . 
DFO and PFN operated an RST during the spring outmigration of 2023 (pilot year), 2024 and 2025. We also conducted purse seining to survey early marine entry during these years. We also include here RST results from 2006 and 2007 to provide a discontinuous time series of data in order to increase our sample size and inference (weird wording). 
We also integrate beach seine monitoring results over x years (range) from PFN’s ongoing monitoring program (PFN and D Burt and Assoc). We are grateful for the data sharing in order to make more robust inference about San Juan juvenile Chinook.  

## Lab processing 

## Data analyses 




# Results

We are so amaaaaaaazing

```{r fig.cap='Daily catch totals from 2023 purse seine. "Other Misc Non-salmon" include incidental catches of juvenile Pacific Sandfish, Anchovy, Shiner Perch, juvenile Kelp Greenling, and Sardine. One crab and one dogfish are also excluded from the figure.'}
PRS_daily_catch_fig_dodge
```

```{r fig.cap='Daily catch totals from 2023 purse seine. "Other Misc Non-salmon" include incidental catches of juvenile Pacific Sandfish, Anchovy, Shiner Perch, juvenile Kelp Greenling, and Sardine. One crab and one dogfish are also excluded from the figure.'}
PRS_daily_catch_fig_stack
```


Some more text about how gr8 we are

```{r fig.cap='Cumulative probability of chinook capture throughout the 2023 purse seine season. Due to exceptionally low catches of natural-origin Chinook, the figure combines both natural and hatchery-origin Chinook.'}
cuml_prob_fig
```



# Discussion











