---
title: "Area 20/San Juan marine juvenile summary (DRAFT)"
author: "South Coast StA"
date: 'Last update: `r format(Sys.time(), sep="_", "%Y-%m-%d %H:%M")`'
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, echo=F, message=F)

# Load libraries ------------------------------------
library(tidyverse)
library(readxl)
library(kableExtra)
library(bookdown)
library(here)
library(scales)       # for label_percent()
library(ggpubr)       # for ggarrange()
library(plotly)       # for ggplotly()



# Helpers ------------------------------------
"%notin%" <- Negate("%in%")



# ===================== LOAD DATA =====================

# J. King WCVI Trawl Data ------------------------------------
swvi.trawl <- read.csv(here("data", "juvenile", "marine", "juvenile_chinook_AOI_20240412.csv")) %>%
  mutate(stock_group = case_when(above_stock_prob_limit_yn=="Y" & grepl("NITINAT", stock_final) ~ "SWVI",
                                 above_stock_prob_limit_yn=="Y" & grepl("JUAN", stock_final) ~ "San Juan",
                                 above_stock_prob_limit_yn=="Y" & grepl("NECHAKO|CLEARWATER|CHILLIWACK|CHEHALIS|THOMPSON", stock_final) ~ "Fraser",
                                 above_stock_prob_limit_yn=="Y" & grepl("WENATCHEE|SOOS|ABERNATHY|NOOKSACK", stock_final) ~ "Southern US",
                                 TRUE ~ "Not grouped"))
```

<br>

## **Early marine entry (summer purse seine)**

**Download raw purse seine data with genetic results from here: https://github.com/khdavidson/A20/blob/main/outputs/R_OUT%20-%20PFN_DFO_FTFjuvi_2023_verified%20WITH%20RESULTS.xlsx**

<br>

### **Catch overview**

The following purse seine events occurred in 2023: 

```{r}
juvi.bioMeta %>%
  filter(grepl("purse seine", gear)) %>% 
  group_by(date_start) %>% 
  summarize() %>% 
  kbl(align="c", caption="2023 purse seining events in Port Renfrew.") %>%
  kable_paper("hover", full_width=T, position = "center")
```

<br>

Most Chinook were encountered in late June, with a later bump in catch in mid-August. Forage fish (herring, smelts) were the most common species, followed by Chinook and then Sandlance (one large catch on July 7). Purse seining likely started too late to encounter other salmon species (Coho, Chum) (see \@ref(fig:purse-catch)). 

```{r include=F}
catch_summary %>% 
  kbl(align="c", caption="2023 purse seining catch composition in Port Renfrew.") %>%
  kable_paper("hover", full_width=T, position = "center")
```

```{r purse-catch, fig.cap='2023 purse seine catch composition off Port Renfrew (Port San Juan). Numbers above bars represent total daily catch across all sets.'}

ggplot() +
  geom_bar(data=catch_summary, aes(x=as.Date(date_start), y=catch_propn, fill=spp_stage, colour=spp_stage), 
           stat="identity", alpha=0.8, width=2) +
  geom_text(data=catch_summary%>%group_by(date_start)%>%summarize(total=unique(daily_total)), 
            aes(x=as.Date(date_start), y=1.05, label=total),
            stat="identity", size=3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_date(date_breaks = "7 day", date_labels="%b %d") +
  labs(x="", y="Proportion of daily total catch", fill="Species/life stage", colour="Species/life stage") +
  theme_bw()
```

<br>

```{r}
# ********** ADD LEAFLET MAP OF CATCH BY DATE/SITE
```





### **Chinook catch and biosamples** (in progress)

Very few adipose-clipped hatchery Chinook were encountered, but clipped fish were generally larger than unclipped fish. However, note that on July 7, all fish with DNA results returned were hatchery-origin San Juan (1 clipped, 20 unclipped), with the exception of 1 unclipped Washington-origin hatchery fish (Willapa River). 

```{r cn-catch-comp, fig.cap='2023 purse seine catch composition in Port Renfrew of chinook only.'}

ggplot(data=cn_mark_summary, aes(x=as.Date(date_start), y=daily_propn, fill=ad_clip, colour=ad_clip, label=n)) +
  geom_bar(stat="identity", width=2, alpha=0.8) +
  geom_text(stat="identity", colour="black", position = position_stack(vjust=0.5)) +
  labs(x="", y="Daily Chinook catch composition", fill="Adipose clip?", colour="Adipose clip?") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_date(date_breaks = "7 day", date_labels="%b %d") +
  theme_bw()
```

<br>

```{r prs-size, fig.cap='Size of Chinook (marked and unmarked) in 2023 purse seine samples in Port Renfrew.'}
ggplot() +
  geom_boxplot(data=juvi.bioMeta%>%filter(species=="chinook", grepl("purse seine", gear)), 
               aes(x=as.Date(date_start), y=length, group=interaction(date_start, ad_clip), fill=ad_clip, colour=ad_clip), 
               stat="boxplot", alpha=0.8, width=3) +
  scale_x_date(date_breaks = "7 day", date_labels = "%b %d") +
  labs(x="", y="Fork length (mm)", fill="Adipose clip?", colour="Adipose clip?") +
  theme_bw()
```



```{r stock-comp-fig, fig.cap='GSI stock comps only from the samples collected July 7. Reminaing samples still in process.'}

ggplot() +
   geom_bar(data=juvi_summary%>%filter(grepl("purse seine", gear)), 
            aes(x=as.Date(date_start), y=day_ID_propn, fill=`(R) STOCK-ORIGIN`), colour="transparent", stat="identity", size=2, alpha=0.6) +
  geom_text(data=juvi_summary%>%filter(grepl("purse seine", gear))%>%arrange(desc(`(R) STOCK-ORIGIN`)), 
            aes(x=as.Date(date_start), y=day_ID_propn, label=n), position = position_stack(vjust=0.5)) +
  geom_bar(data=juvi_summary%>%filter(grepl("purse seine", gear)) %>% group_by(date_start, `(R) ORIGIN`) %>% summarize(day_origin_propn=unique(day_origin_propn)), 
           aes(x=as.Date(date_start), y=day_origin_propn, colour=`(R) ORIGIN`), fill="transparent", stat="identity", size=2) +
  scale_colour_manual(breaks=waiver(), values = c("black", "transparent")) +
  scale_y_continuous(labels = scales::percent)  +
  scale_x_date(date_labels = "%b %d") +
  labs(x="", y="Daily stock composition") +
  theme_bw()
```


<br>

## !!!  Outstanding genetic samples 2023 !!! 

GSI results recently returned were only available for July 7th purse seining. Still waiting on the remaining days: 

```{r}
# Sample processing overview: samples analyzed and remaining to analyze
genetic.samples23 <- juvi.bioMeta %>% 
  filter(grepl("purse seine|beach seine", gear), !is.na(DNA_vial)) %>%
  group_by(gear, lubridate::month(date_start), date_start, DNA_vial, is.na(indiv)) %>% 
  summarize(n=n()) %>% 
  mutate(n = case_when(`is.na(indiv)`=="TRUE" ~ 0,
                       TRUE ~ n)) %>% 
  group_by(gear, date_start) %>% 
  mutate(results_by_date=sum(n)) %>% 
  arrange(date_start) %>% 
  #filter(results_by_date<1) %>% 
  select(-c(results_by_date, `is.na(indiv)`, `lubridate::month(date_start)`)) %>%
  rename(method=gear,
         `date collected`=date_start,
         `DNA vial` = DNA_vial,
         `Results returned? 1=yes, 0=no`=n) %>% 
  mutate_at("date collected", as.Date)

writexl::write_xlsx(genetic.samples23, here("outputs", "San Juan DNA samples outstanding 2023.xlsx"))

genetic.samples23 %>%
  kbl()
```

<br>

<br>

## First marine winter (trawl/microtrolling)

```{r}
ggplotly(
  ggplot() +
    geom_sf(data=rnaturalearth::ne_countries(type="countries", scale="large", returnclass="sf"), 
            colour="transparent", fill="gray60") +
    geom_jitter(data=swvi.trawl%>%filter(above_stock_prob_limit_yn=="Y"), 
                aes(x=longitude, y=latitude, fill=stock_group, colour=stock_group, 
                    text=paste0("Stock:", stock_final)), 
                position=position_jitter(width=0.1), shape=21, size=3, alpha=0.5, stroke=0.1) +
    geom_jitter(data=swvi.trawl%>%filter(above_stock_prob_limit_yn!="Y"), 
                aes(x=longitude, y=latitude, text=paste0("Stock:", stock_final), fill=stock_group, colour=stock_group), 
                position=position_jitter(width=0.1), shape=21, size=3, alpha=0.2, stroke=0.1) +
    coord_sf(crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0", xlim=c(-125.5, -123.5), ylim=c(48.8, 48.2)) +
    theme_bw() ,
  tooltip = c("text")
)


```
