---
title: "part4-juvi-knitable"
author: "South Coast StA"
date: 'Last update: `r format(Sys.time(), sep="_", "%Y-%m-%d %H:%M")`'
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, echo=F, message=F)

# Load libraries ------------------------------------
library(tidyverse)
library(readxl)
library(kableExtra)
library(bookdown)
library(here)
library(scales)       # for label_percent()
library(ggpubr)       # for ggarrange()



# Helpers ------------------------------------
"%notin%" <- Negate("%in%")







# ===================== LOAD DATA =====================


# San Juan juvi Field Database ------------------------------------
juvi.meta <- readxl::read_excel(here("data", "juvenile", "PFN_DFO_FTFjuvi_2023_verified.xlsx"), sheet="sample_event_metadata", trim_ws=T)
juvi.enocgy <- readxl::read_excel(here("data", "juvenile", "PFN_DFO_FTFjuvi_2023_verified.xlsx"), sheet="enviro_ocgy", trim_ws=T)
juvi.totals <- readxl::read_excel(here("data", "juvenile", "PFN_DFO_FTFjuvi_2023_verified.xlsx"), sheet="set_totals", trim_ws=T)
juvi.bio <- readxl::read_excel(here("data", "juvenile", "PFN_DFO_FTFjuvi_2023_verified.xlsx"), sheet="biosamples", trim_ws=T) 
                      
  

# WCVI juvenile data ------------------------------------
wcvi.juviDS <- readxl::read_excel(path="//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/JUVENILE_PROJECTS/DOWNSTREAM/WCVI Juvenile Down Stream Data Set 1972-present.xlsx", sheet="DS Summary", skip=2) %>% 
  filter(`Data Type` != "Total Count") %>%
  mutate(date = case_when(!is.na(Day) ~ lubridate::ymd(paste0(Year, substr(Day, 6, 10), sep="-")),
                          TRUE ~ NA),
         yday = lubridate::yday(date)) %>%
  pivot_longer(c(`Coho smolt`:Chum, Steelhead), names_to = "Species", values_to = "Count") %>% 
  group_by(Year, date, yday, `Data Type`, `Stat Area`, Location, Type, Species, `Temp (C)`) %>% 
  summarize(Count=sum(Count))


# Historical SJ ------------------------------------
SJ06 <- readxl::read_excel(path="//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/JUVENILE_PROJECTS/DOWNSTREAM/SAN JUAN RST/2006/SanJuanRSTReport/working files and data/2006-SanJuan-frybiosampling.xls", sheet="Data") %>% 
  mutate(yday = lubridate::yday(date))



# 2023 GSI ------------------------------------
# Also links juvi.meta, juvi.bio and MGL outputs
source(here("scripts","misc-helpers","gsiCompile.R")) 









# ===================== JOIN DATA =====================

# Field catch totals + meta ------------------------------------
juvi.totalMeta <- left_join(juvi.totals,
                            juvi.meta,
                            by="usid") %>% 
  unite(spp_stage, c(species, life_stage), remove = F, na.rm = T, sep=" ") %>% 
  mutate(spp_group = case_when(species%in%c("coho","chinook","chum","pink","sockeye") ~ "salmon",
                             TRUE ~ "non-salmonid")) %>% 
  print()








# ===================== SUMMARIZE DATA =====================
# GSI RESULTS SUMMARY ------------------------------------
juvi_summary <- juvi.bioMeta %>% 
  filter(!grepl("dip net|pole seine", gear), !is.na(DNA_vial), !is.na(`(R) STOCK ID`)) %>% 
  group_by(date_start, gear, `(R) ORIGIN`, `(R) STOCK-ORIGIN`) %>% 
  summarize(n=n()) %>% 
  group_by(date_start, `(R) ORIGIN`) %>%
  mutate(origin_total=sum(n)) %>% 
  group_by(date_start) %>%
  mutate(day_total=sum(n),
         day_ID_propn = n/day_total,
         day_origin_propn = origin_total/day_total) %>% 
  print()


# Catch summary ------------------------------------
catch_summary <- juvi.totalMeta %>% 
  filter(grepl("purse seine", gear)) %>% 
  group_by(date_start, spp_group, spp_stage) %>% 
  summarize(n = sum(total_caught)) %>% 
  group_by(date_start) %>% 
  mutate(daily_total = sum(n),
         catch_propn = n/daily_total) %>%
  print()


# Chinook mark/unmarked ------------------------------------
cn_mark_summary <- juvi.bioMeta %>% 
  filter(grepl("purse seine", gear), species=="chinook") %>% 
  group_by(date_start, ad_clip) %>% 
  summarize(n=n()) %>% 
  group_by (date_start) %>% 
  mutate(daily_total = sum(n),
         daily_propn = n/daily_total) %>% 
  print()
```

# **San Juan juveniles**

## **Freshwater - RST**

**Download raw RST data with genetic results from here: https://github.com/khdavidson/A20/blob/main/outputs/R_OUT%20-%20PFN_DFO_FTFjuvi_2023_verified%20WITH%20RESULTS.xlsx** 

<br>

### **Catch overview**

<br>

Coho fry were by far the most common species encountered in the San Juan RST in 2023. Discerning any outmigration patterns is difficult from the 2023 data as it was only run 2 days/week, but suggests an earlier install may be warranted in future years (see \@ref(fig:rst-total)). 

```{r rst-total, fig.cap='Total catch in the RST by salmon species and life stage. By-catch of trout and other non-fish species not shown.'}
ggplot() +
  geom_bar(data=juvi.totalMeta%>%
             filter(grepl("RST", gear), grepl("coho|chum|chinook",spp_stage), life_stage!="alevin") %>%
             group_by(date_end, spp_stage) %>% 
             summarize(n=sum(total_caught)),
           aes(x=as.Date(date_end), y=n, fill=spp_stage, colour=spp_stage), stat="identity", alpha=0.8) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d") +
  labs(x="",y = "Catch", fill="Species/life stage", colour="Species/life stage") +
  theme_bw() +
  theme(axis.text = element_text(colour="black"),
        axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_text(face="bold"),
        legend.title = element_text(face="bold"))
```


<br>

### **Chinook catch and biosamples**

<br>

In 2023 we caught relatively few Chinook in the RST. This is likely due to the trap only running 2 days/week as a pilot project. 2006 and 2007 catch data are given below, as well as Bedwell RST chinook catch, in order to compare timing and relative magnitude (see \@ref(fig:rst-rel)). Note that San Juan 2006 saw exceptionally high catch of chinook in one night, but was otherwise relatively consistent. 

While the RST has had a fairly limited season at San Juan (no early February install or late June removal as at Sarita), the migration timing generally follows other WCVI systems, with many small fish leaving ~March-April. 

```{r rst-rel}
ggplot() +
  geom_bar(data=juvi.totalMeta%>%filter(grepl("RST", gear), grepl("chinook",spp_stage), life_stage!="alevin")%>%
             group_by(date_end, spp_stage)%>%summarize(n=sum(total_caught)),
           aes(x=as.Date(date_end), y=n, fill=spp_stage, colour=spp_stage), stat="identity", alpha=0.8) +
  labs(x="",y="Number of chinook caught") +
  theme_bw() +
  theme(legend.position = "none")



ggplot() +
  geom_bar(data=wcvi.juviDS%>%filter(Species=="Chinook", Location%in%c("San Juan River", "Bedwell River")),
           aes(x=as.Date(yday, origin=as.Date("1992-12-31")), y=Count, 
               group=interaction(Year,Location), fill=as.factor(Year)), 
           stat="identity", position="identity",  alpha=0.5, size=0, width=1) +
  scale_x_date(date_breaks = "7 day", date_labels="%b %d") +
  labs(x="",y="Number of chinook", fill="", colour="") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text= element_text(size=12, colour="black"),
        axis.title = element_text(size=15, face="bold"),
        legend.text = element_text(size=10)) +
  facet_wrap(~Location, scales="free_y", nrow=2)
```

<br>

Body size of San Juan chinook follow similar trends to other locations (Bedwell, Sarita), with early migrants being very small (mean FL = 40.9 +/- 2.21 mm and 0.57 +/- 0.08 g). The weight of natural San Juan Chinook is exceptionally small compared to the size of hatchery releases, which are typically  5-8 grams; even historical releases of fed fry were 1.5-2.5 g (see \@ref(fig:sizerel-fig)). 

```{r}
ggplot() +
    geom_boxplot(data=SJ06%>%filter(`hatchery code`=="wild", species==124),
             aes(x=as.Date(yday, origin=as.Date("1970-12-31")), y=length, group=date), fill="gray80", alpha=0.8) +
  
    geom_boxplot(data=juvi.bioMeta%>%filter(grepl("RST",gear), species=="chinook"),
             aes(x=as.Date(yday, origin=as.Date("1970-12-31")), y=length, group=date_end), 
             fill="dodger blue", width=2.5, alpha=0.8) +
  scale_y_continuous(limits=c(20,60)) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d", limits = c(as.Date(69, origin=as.Date("1970-12-31")), 
                                                                    as.Date(205, origin=as.Date("1970-12-31")))) +
  labs(x="", y="Mean fork length (mm)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(size=12, colour="black"),
        axis.title = element_text(size=15, face="bold")) 



ggplot() +
    geom_boxplot(data=juvi.bioMeta%>%filter(grepl("RST",gear), species=="chinook"),
             aes(x=as.Date(yday, origin=as.Date("1970-12-31")), y=weight, group=date_end), 
             fill="dodger blue", width=2.5, alpha=0.8) +
  scale_y_continuous(limits=c(0,1.5)) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d", limits = c(as.Date(69, origin=as.Date("1970-12-31")), 
                                                                    as.Date(205, origin=as.Date("1970-12-31")))) +
  labs(x="", y="Mean weight (g)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(size=12, colour="black"),
        axis.title = element_text(size=15, face="bold")) 
```


```{r}
juvi.bioMeta %>%
  filter(grepl("RST", gear), species=="chinook") %>% 
  summarize(meanFL=mean(length), sdFL=sd(length),
            meanW=mean(weight), sdW=sd(weight))
```

<br>

Only a small number of genetic samples were collected from the RST as we figured we did not need to bother running genetics from known San Juan origin juveniles upstream of the hatchery release location. However, of the 7 samples collected on one day, 3 were inconclusive/non-San Juan!

```{r rst-comps, fig.cap='Genetic results from the San Juan RST 2003. Samples were not collected as they were not considered a focus of the project.'}
ggplot() +
   geom_bar(data=juvi_summary%>%filter(grepl("RST", gear)), 
            aes(x=as.Date(date_start), y=day_ID_propn, fill=`(R) STOCK-ORIGIN`), colour="transparent", stat="identity", size=2, alpha=0.6) +
  geom_text(data=juvi_summary%>%filter(grepl("RST", gear))%>%arrange(desc(`(R) STOCK-ORIGIN`)), 
            aes(x=as.Date(date_start), y=day_ID_propn, label=n), position = position_stack(vjust=0.5)) +
  geom_bar(data=juvi_summary%>%filter(grepl("RST", gear)) %>% group_by(date_start, `(R) ORIGIN`) %>%
             summarize(day_origin_propn=unique(day_origin_propn)), 
           aes(x=as.Date(date_start), y=day_origin_propn, colour=`(R) ORIGIN`), fill="transparent", stat="identity", size=2) +
  scale_colour_manual(breaks=waiver(), values = c("white", "black", "transparent")) +
  scale_y_continuous(labels = scales::percent)  +
  scale_x_date(date_labels = "%b %d") +
  labs(x="", y="Daily stock composition") +
  theme_bw()
```



<br>

<br>

## **Estuary (PFN beach seine)**

<br>

(TBD - require updated PFN beach seine data/report)

<br>

## **Early marine entry (summer purse seine)**

**Download raw purse seine data with genetic results from here: https://github.com/khdavidson/A20/blob/main/outputs/R_OUT%20-%20PFN_DFO_FTFjuvi_2023_verified%20WITH%20RESULTS.xlsx**

<br>

### **Catch overview**

The following purse seine events occurred in 2023: 

```{r}
juvi.bioMeta %>%
  filter(grepl("purse seine", gear)) %>% 
  group_by(date_start) %>% 
  summarize() %>% 
  kbl(align="c", caption="2023 purse seining events in Port Renfrew.") %>%
  kable_paper("hover", full_width=T, position = "center")
```

<br>

Most Chinook were encountered in late June, with a later bump in catch in mid-August. Forage fish (herring, smelts) were the most common species, followed by Chinook and then Sandlance (one large catch on July 7). Purse seining likely started too late to encounter other salmon species (Coho, Chum) (see \@ref(fig:purse-catch)). 

```{r include=F}
catch_summary %>% 
  kbl(align="c", caption="2023 purse seining catch composition in Port Renfrew.") %>%
  kable_paper("hover", full_width=T, position = "center")
```

```{r purse-catch, fig.cap='2023 purse seine catch composition off Port Renfrew (Port San Juan). Numbers above bars represent total daily catch across all sets.'}

ggplot() +
  geom_bar(data=catch_summary, aes(x=as.Date(date_start), y=catch_propn, fill=spp_stage, colour=spp_stage), 
           stat="identity", alpha=0.8, width=2) +
  geom_text(data=catch_summary%>%group_by(date_start)%>%summarize(total=unique(daily_total)), 
            aes(x=as.Date(date_start), y=1.05, label=total),
            stat="identity", size=3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_date(date_breaks = "7 day", date_labels="%b %d") +
  labs(x="", y="Proportion of daily total catch", fill="Species/life stage", colour="Species/life stage") +
  theme_bw()
```

<br>





### **Chinook catch and biosamples** (in progress)

Very few adipose-clipped hatchery Chinook were encountered, but clipped fish were generally larger than unclipped fish. However, note that on July 7, all fish with DNA results returned were hatchery-origin San Juan (1 clipped, 20 unclipped), with the exception of 1 unclipped Washington-origin hatchery fish (Willapa River). 

```{r cn-catch-comp, fig.cap='2023 purse seine catch composition in Port Renfrew of chinook only.'}

ggplot(data=cn_mark_summary, aes(x=as.Date(date_start), y=daily_propn, fill=ad_clip, colour=ad_clip, label=n)) +
  geom_bar(stat="identity", width=2, alpha=0.8) +
  geom_text(stat="identity", colour="black", position = position_stack(vjust=0.5)) +
  labs(x="", y="Daily Chinook catch composition", fill="Adipose clip?", colour="Adipose clip?") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_date(date_breaks = "7 day", date_labels="%b %d") +
  theme_bw()
```

<br>

```{r prs-size, fig.cap='Size of Chinook (marked and unmarked) in 2023 purse seine samples in Port Renfrew.'}
ggplot() +
  geom_boxplot(data=juvi.bioMeta%>%filter(species=="chinook", grepl("purse seine", gear)), 
               aes(x=as.Date(date_start), y=length, group=interaction(date_start, ad_clip), fill=ad_clip, colour=ad_clip), 
               stat="boxplot", alpha=0.8, width=3) +
  scale_x_date(date_breaks = "7 day", date_labels = "%b %d") +
  labs(x="", y="Fork length (mm)", fill="Adipose clip?", colour="Adipose clip?") +
  theme_bw()
```









```{r stock-comp-fig, fig.cap='GSI stock comps only from the samples collected July 7. Reminaing samples still in process.'}

ggplot() +
   geom_bar(data=juvi_summary%>%filter(grepl("purse seine", gear)), 
            aes(x=as.Date(date_start), y=day_ID_propn, fill=`(R) STOCK-ORIGIN`), colour="transparent", stat="identity", size=2, alpha=0.6) +
  geom_text(data=juvi_summary%>%filter(grepl("purse seine", gear))%>%arrange(desc(`(R) STOCK-ORIGIN`)), 
            aes(x=as.Date(date_start), y=day_ID_propn, label=n), position = position_stack(vjust=0.5)) +
  geom_bar(data=juvi_summary%>%filter(grepl("purse seine", gear)) %>% group_by(date_start, `(R) ORIGIN`) %>% summarize(day_origin_propn=unique(day_origin_propn)), 
           aes(x=as.Date(date_start), y=day_origin_propn, colour=`(R) ORIGIN`), fill="transparent", stat="identity", size=2) +
  scale_colour_manual(breaks=waiver(), values = c("black", "transparent")) +
  scale_y_continuous(labels = scales::percent)  +
  scale_x_date(date_labels = "%b %d") +
  labs(x="", y="Daily stock composition") +
  theme_bw()
```





<br>

## !!!  Outstanding genetic samples !!! 

GSI results recently returned were only available for July 7th purse seining. Still waiting on the remaining days: 

```{r}
# Sample processing overview: samples analyzed and remaining to analyze
juvi.bioMeta %>% 
  filter(grepl("purse seine|beach seine", gear), !is.na(DNA_vial)) %>%
  group_by(gear, lubridate::month(date_start), date_start, DNA_vial, is.na(indiv)) %>% 
  summarize(n=n()) %>% 
  mutate(n = case_when(`is.na(indiv)`=="TRUE" ~ 0,
                       TRUE ~ n)) %>% 
  group_by(gear, date_start) %>% 
  mutate(results_by_date=sum(n)) %>% 
  arrange(date_start) %>% 
  #filter(results_by_date<1) %>% 
  select(-c(results_by_date, `is.na(indiv)`, `lubridate::month(date_start)`)) %>%
  rename(method=gear,
         `date collected`=date_start,
         `DNA vial` = DNA_vial,
         `Results returned? 1=yes, 0=no`=n) %>% 
  kbl()
```

<br>




<br>

<br>












