---
title: "**Part 5: Environmental details**"
date: 'Last update: `r format(Sys.time(), sep="_", "%Y-%m-%d %H:%M")`'
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
params:
    password:
      label: "Enter computer password"
      value: ""
      input: password
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)

#library(here)
library(tidyverse)
#library(readxl)
#library(kableExtra)
library(bookdown)
#library(here)

"%notin%" <- Negate("%in%")


# ====================== LOAD DATA ======================
# Environment Canada historical water data (1959-2022): --------------------------
ec.hist <- read.csv(file=paste0(here::here("data", "enviro","/"), 
                                list.files(path=here::here("data", "enviro"), pattern="08HA010_daily*.*csv")), 
                    skip=1) %>% 
  mutate(discharge_cms = case_when(PARAM==1 ~ Value),
         level_m = case_when(PARAM==2 ~ Value),
         year = lubridate::year(Date),
         Date = as.Date(Date),
         DOY = lubridate::yday(Date),
         PARAM = case_when(PARAM==1 ~ "discharge",
                           PARAM==2 ~ "level",
                           TRUE~"FLAG")) %>% 
  print()


# Environment Canada real-time water data (2023): --------------------------
ec.rt <- read.csv(url("https://dd.weather.gc.ca/hydrometric/csv/BC/daily/BC_08HA010_daily_hydrometric.csv")) %>% 
  rename(level_m=`Water.Level...Niveau.d.eau..m.`,
         discharge_cms = `Discharge...Débit..cms.`) %>% 
  mutate(Date = as.Date(Date)) %>%
  print()


# Fence install dates -------------------------- 
sj.fence.dates <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/run timing tables (updated).xlsx", sheet="San Juan", range="A4:AE65") %>% 
  rename(year=`...1`,
         system=`81`) %>%
  fill(year) %>%
  filter(year!=max(year) | !is.na(system), !is.na(`Fence in`)) %>%
  select(year, `Fence in`, `Fence out`) %>%
  mutate(across(c(`Fence in`, `Fence out`), as.Date),
         In = lubridate::yday(`Fence in`),
         Out = lubridate::yday(`Fence out`)) %>% 
  #pivot_longer(cols=c(In, Out), names_to = "install", values_to = "date") %>%
  print()
```

<br>

<br>

## **4 Mile fence operations and flow**

<br> 

The fence operated by 4 Mile hatchery primarily functions to collect broodstock, and secondarily to count all species through. In recent yeras, the operation period is typically from Labour Day long weekend to Thanksgiving long weekend, as flow allows. It is usually removed just in time for the fall/winter floods (see Fig \@ref(fig:fenceflowfig)), around approximately 5 cms or 2 m, but can be as high as 35 cms (see Table \@ref(tab:removalvalstab)). 

```{r fenceflowfig, fig.cap='4 Mile fence operation periods (red segments) and observed discharge (dark blue, cms) and water level (light blue, m, where available). Data from 4 Mile hatchery and Environment Canada.'}
# Fence dates ~ discharge

ggplot() +
  geom_line(data=ec.hist%>%filter(year %in% sj.fence.dates$year, year%notin%c(1995,1996), !is.na(discharge_cms)), 
            aes(x=as.Date(DOY, origin="1995-12-31"), y=discharge_cms, group=interaction(year)),
            size=1, colour="navy blue") +
    geom_line(data=ec.hist%>%filter(year %in% sj.fence.dates$year, year%notin%c(1995,1996), !is.na(level_m)), 
            aes(x=as.Date(DOY, origin="1995-12-31"), y=level_m*100, group=interaction(year)),
            size=1, colour="dodger blue") +
  geom_segment(data=sj.fence.dates%>%filter(year%notin%c(1995,1996)), 
               aes(x=as.Date(In, origin="1995-12-31"), y=100, xend=as.Date(Out, origin="1995-12-31"), yend=100), colour="red", size=2)  +
  scale_colour_manual(breaks=c(1,2), values = c("navy blue", "dodger blue")) +
  scale_x_date(date_labels="%b %d", limits = c(as.Date(230, origin="1995-12-31"), as.Date(350, "1995-12-31"))) +
  scale_y_continuous(sec.axis = sec_axis(~ . /100, name = "Level (m)")) +
  labs(x="", y="Discharge (cms)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_wrap(~year)
```

<br>

```{r removalvalstab}
ec.hist %>% 
  filter(Date %in% sj.fence.dates$`Fence out`) %>% 
    group_by(PARAM) %>% 
    summarize(min_val = min(Value, na.rm=T),
              max_val = max(Value, na.rm=T),
              med_val = median(Value, na.rm=T)) %>% 
  relocate(med_val, .before=max_val) %>%
  rename(Parameter=PARAM,
         `Minimum value` = min_val,
         `Maximum value` = max_val,
         `Median value` = med_val)  %>%
  kableExtra::kbl(align="c", caption="Range and median river discharge (cms) and/or level (m) for known fence removal dates illustrated in Figure 1 above.") %>% 
  kableExtra::kable_paper("hover", full_width=T, position = "center")

# figure below - probably not required. 
```
```{r include=F, eval=F}
# Fence removal date vs. water fig - same as table above, probably not needed
ggplot() +
  geom_point(data=ec.hist %>% filter(Date %in% sj.fence.dates$`Fence out`),
             aes(x=year, y=discharge_cms), colour="navy blue", fill="navy blue", size=3, shape=21) +
  geom_point(data=ec.hist %>% filter(Date %in% sj.fence.dates$`Fence out`),
             aes(x=year, y=level_m*5), colour="dodger blue", fill="dodger blue", size=3, shape=21) +
  scale_x_continuous(breaks=seq(1997, lubridate::year(Sys.Date()), by=2)) +
  scale_y_continuous(sec.axis = sec_axis(~ . /5, name = "Level (m)")) +
  labs(y="Discharge (cms)", x="") +
  theme_bw()
```















